---
title: "Morphospace"
output: html_document
date: "2025-09-26"
---
```{r}
library(survival)
library(tidyverse)
library(finalfit)
library(dplyr)
library(readxl)
library(readr)
library(forcats)
library(beepr)
library(cowplot)
library(tidyr)
library(scales)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(stringr)
library(splitstackshape)
library(MASS)
library(investr)
library(survminer)
#detach("package:plyr", unload = TRUE)
setwd("~/Documents/SSD_Data") #folder files are in

L_col <- "#004c84"  # Low line color
L.col <- '#004c84'
H_col <- "#ffb656"  # High line color
H.col <- '#ffb656'
C_col <- "#7f7f7f"
C.col <- '#7f7f7f'

iVf <- read.csv("initialVfinalFamily.csv")
diffdat <- read_excel("SSD_means_.xlsx", sheet="generational differentials")
rawfvw <- read_excel("tonyFVWraw.xlsx", sheet = "COMBINE")
```
```{r filtering FVW}
fvw <- drop_na(rawfvw)
fvw$logpupa <- log(fvw$pupa)
fvw$logwing <- log(fvw$wing)

fvwtop <- fvw %>%
  group_by(lineage, sex) %>%
#  filter(pupa < quantile(pupa, 0.15) | pupa > quantile(pupa, 0.85)) %>%
  filter(logpupa > quantile(logpupa, 0.75)) %>%
    ungroup()

head(fvwtop)
```

```{r initial vs final}
intVfin <- ggplot(iVf, aes(x=mean.m, y=mean.f, group=line, color=line)) +
  geom_point(aes(shape = line, size = .1, alpha = line)) +
  scale_shape_manual(values = c("C" = 16, "H" = 16, "I" = 1, "L" = 16)) +
  scale_color_manual(values = c("C" = C_col, 
                               "H" = H_col, 
                               "I" = "black", 
                               "L" = L_col)) +
  scale_alpha_manual(values = c("C" = 0.5, "H" = 0.5, "I" = 1, "L" = 0.5), guide = "none") + # Consistent alpha
  theme_cowplot()

intVfin
#ggsave("C:/Users/lizch/Desktop/Data_Art_SSD/Committee2025/intVfin.png", width = plot_width/2, height = plot_height, bg = "white")
#ggsave("C:/Users/lizch/Desktop/Data_Art_SSD/Committee2025/intVfin.svg", width = plot_width/2, height = plot_height, bg = "white")

#subset data
FvML <- subset(diffdat, select = c(generation, L.mean.f, L.mean.m))
FvMH <- subset(diffdat, select = c(generation, H.mean.f, H.mean.m))
FvMC <- subset(diffdat, select = c(generation, C.mean.f, C.mean.m))

#add line
FvML$line <- as.factor("L")
FvMH$line <- as.factor("H")
FvMC$line <- as.factor("C")

#rename columns
names(FvML)[names(FvML) == "L.mean.f"] <- "mean.f"
names(FvML)[names(FvML) == "L.mean.m"] <- "mean.m"
names(FvMH)[names(FvMH) == "H.mean.f"] <- "mean.f"
names(FvMH)[names(FvMH) == "H.mean.m"] <- "mean.m"
names(FvMC)[names(FvMC) == "C.mean.f"] <- "mean.f"
names(FvMC)[names(FvMC) == "C.mean.m"] <- "mean.m"
FvM <- rbind(FvMC, FvMH, FvML)
print(FvM)

FvMplot <- ggplot(FvM, aes(x = mean.m, y = mean.f, group=line, color=line)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "black") +
  scale_y_continuous(limits=c((14.975-.10), 14.975), expand = c(0, 0)) +
  scale_x_continuous(limits=c(14.725, (14.725+.10)), expand = c(0, 0))+
  geom_path(arrow = arrow(angle = 10, ends= "both", type = "closed"),size=0.5)

FvMplot
ggsave("~/Documents/SSD_Data/Committee2025/FvM.png", width = 5, height = 5, bg = "white")
```

```{r}
iVf$dataset <- "points"
FvM$dataset <- "path"

combined_data <- bind_rows(iVf, FvM)

path_arrows <- FvM %>%
  arrange(line, generation) %>%
  group_by(line) %>%
  mutate(
    x_end = lead(mean.m),
    y_end = lead(mean.f),
    generation_end = lead(generation)
  ) %>%
  filter(!is.na(x_end)) %>%  # Remove the last point which has no next generation
  ungroup()

plotstuff <- ggplot(combined_data, aes(x = mean.m, y = mean.f, group = line, color = line)) +
  scale_color_manual(values = c("C" = C_col, "H" = H_col, "I" = "black", "L" = L_col)) +
  theme_cowplot() +
  theme(legend.position = "none")

iVf_points <- plotstuff +
  geom_smooth(data = subset(combined_data, dataset == "points" & line == "I"), method = "glm", color = "black",  linetype = "dashed", se = FALSE, size = 0.5) +
  geom_smooth(data = subset(combined_data, dataset == "points" & line == "C"), method = "glm", color = C_col,  linetype = "dashed", se = FALSE, size = 0.5) +
  geom_smooth(data = subset(combined_data, dataset == "points" & line == "H"), method = "glm", color = H_col,  linetype = "dashed", se = FALSE, size = 0.5) +
  geom_smooth(data = subset(combined_data, dataset == "points" & line == "L"), method = "glm", color = L_col,  linetype = "dashed", se = FALSE, size = 0.5) +
  geom_point(data = subset(combined_data, dataset == "points"), aes(shape = line, alpha = line)) +
  scale_shape_manual(values = c("C" = 16, "H" = 16, "I" = 1, "L" = 16))

iVf_lines <- plotstuff +
  geom_smooth(data = subset(combined_data, dataset == "points" & line == "I"), method = "glm", color = "black",  linetype = "dashed", se = FALSE, size = 0.5) +
  geom_smooth(data = subset(combined_data, dataset == "points" & line == "C"), method = "glm", color = C_col,  linetype = "dashed", se = FALSE, size = 0.5) +
  geom_segment(data = path_arrows,
               aes(x = mean.m, y = mean.f, 
                   xend = x_end, yend = y_end,
                   group = line, color = line),
               arrow = arrow(angle = 10, length = unit(0.1, "inches")),
               size = 0.5) #+
  #geom_abline(intercept is observed SSD slope is 1)

#add ssd is 1:1 line 
#major axis regression instead of glm

iVf_points
iVf_lines
```

iVf_combined <- plotstuff +
  geom_smooth(data = subset(combined_data, dataset == "points" & line == "I"), method = "glm", color = "black",  linetype = "dashed", se = FALSE, size = 0.5) +
  geom_smooth(data = subset(combined_data, dataset == "points" & line == "C"), method = "glm", color = C_col,  linetype = "dashed", se = FALSE, size = 0.5) +
  geom_smooth(data = subset(combined_data, dataset == "points" & line == "H"), method = "glm", color = H_col,  linetype = "dashed", se = FALSE, size = 0.5) +
  geom_smooth(data = subset(combined_data, dataset == "points" & line == "L"), method = "glm", color = L_col,  linetype = "dashed", se = FALSE, size = 0.5) +
  geom_point(data = subset(combined_data, dataset == "points"), aes(shape = line, alpha = line)) +
  scale_shape_manual(values = c("C" = 16, "H" = 16, "I" = 1, "L" = 16)) +
  scale_alpha_manual(values = c("C" = 0.5, "H" = 0.5, "I" = 1, "L" = 0.5), guide = "none") +
  geom_segment(data = path_arrows,
               aes(x = mean.m, y = mean.f, 
                   xend = x_end, yend = y_end,
                   group = line, color = line),
               arrow = arrow(angle = 10, length = unit(0.1, "inches")),
               size = 0.5) +

iVf_combined
ggsave("~/Documents/SSD_Data/Committee2025/FvM.png", width = 5, height = 5, bg = "white")
```
